Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-03-17 paper.tex
Working on 6

If the expression is atomic, then we simply need to construct a new instance of
@VS@ and return it to the @continue@ function.

At the end, the initial invocation to @aeval@ is passed with an indentify
function as halting continuation.
\begin{lstlisting}
def analyze(e: Expr) =
  aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), (ans => ans))
\end{lstlisting}

@@

Based on the refunctionalized AAM from the last section, we use the same @nd@ nondeterministic operator
but further extend it to a fixed-point cached evaluation in this section.
The fixed-point caching algorithm performs a sound over-approximation of all possible
concrete reachable pathes and values, and also prevents non-termination when analyzing diverged programs.

\textbf{Configuration.}
Now we do not have the concept of explicit stack and state,
but to represent the cache, we introduce the configuration @Config@ as
a state-like definition which borrows the components from state but without
continuations.
\begin{lstlisting}
case class Config(e: Expr, env: Env, store: BStore, time: Time)
\end{lstlisting}

\textbf{Cache.}
Recall that the latent assumption of the worklist algorithm in classical AAM
is if we have seen a state @s@, then it means we also have seen the successors
of state @s@.
The caching algorithm we adopted here replay the same assumption but in a big-step
manner: if we have seen the configuration @c@, then it means we also have seen
the values that are evaluated by the configuration @c@.
Here we will apply the fixed-point caching algorithm as described from
\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.

@@

The @Cache@ contains two maps @in@ and @out@ which are both mapping from
configurations to sets of @VS@. The @in@ cache contains mappings from
last iteration of evaluation, and @out@ cache contains mappings after this
iteration of evaluation. Once we have evaluated a term to some values,
we update the @out@ cache. In the next time of iteration, we will use @out@
from last iteration as @in@ cache for this iteration.

\textbf{Cache-Passing Style.}
Now we can modify out interpreter to cache-passing style.
First the return type @Ans@ is changed to a case class that contains a set of @VS@ and
a cache.

@@

The skeleton of function @aeval@ largely remain the same.
Every place has a value of @Ans@ is also replaced as direct pattern matching so
that we can use the fields in @Ans@ immediately.
The cache is used in a monotonic way: each function call to @aeval@ or @nd@ is
passed with the latest cache from the most recent continuation's result.

When we enter to function @aeval@, we firstly lookup in the @out@ cache,
if there are some values for current configuration, then we use immediately.
This corresponds to small-step AAM with a worklist when we have seen a state
then we can simply discard the state and continue working the rest of worklist @todo@.

Otherwise, if we not hit the @out@ cache, we retrieve the values from @in@ cache, and update @out@ cache
with values from @in@ cache.  The retrieving from @in@ cache may return an empty set of values if
there is no such mapping for the configuration from the previous iteration.
It actually first conservatively assumes that all computation can be diverged,
if not, we will update its values in the mapping after evaluation.
In terms of partial orders and lattices, it is the case that starts
the Kleene iteration from the bottom of the lattice.

>>>>

If the expression is atomic, then we simply construct a new instance of
@VS@ and return it to the @continue@ function.

At the end, the initial invocation to @aeval@ is passed with an identity
function as a halting continuation:
\begin{lstlisting}
def analyze(e: Expr) =
  aeval(e, Map[String, BAddr](), Store[BAddr, Storable](Map()), List(), (ans => ans))
\end{lstlisting}

@@

Based on the refunctionalized AAM from the last section, we use the same nondeterministic
operator @nd@, but further extend it to a fixed-point cached evaluation in this section.
The fixed-point caching algorithm performs a sound over-approximation of all possible
concrete reachable paths and values, and also prevents non-termination when analyzing diverged programs.

\textbf{Configuration.}
We do not have the concept of an explicit stack or state,
so to represent the cache we introduce a configuration @Config@ as
a state-like definition which borrows the components from state, sans continuations.
\begin{lstlisting}
case class Config(e: Expr, env: Env, store: BStore, time: Time)
\end{lstlisting}

\textbf{Cache.}
Recall that the latent assumption of the worklist algorithm in classical AAM
is that if we have seen a state @s@, it means we also have seen the successors
of state @s@.
The caching algorithm we adopt here replays the same assumption but in a big-step
manner: if we have seen the configuration @c@, then it means we also have seen
the values that are evaluated by the configuration @c@.
Here we will apply the fixed-point caching algorithm as described from
\citeauthor{darais2017abstracting}'s ADI paper \cite{darais2017abstracting}.

@@

The @Cache@ contains two maps @in@ and @out@ which are both mappings from
configurations to sets of @VS@. The @in@ cache contains mappings from the
previous iteration of evaluation, and the @out@ cache contains mappings after the current
iteration of evaluation. Once we have evaluated a term to some values,
we update the @out@ cache. In the next iteration, we will use the @out@
from the previous iteration as the @in@ cache for the current iteration.

\textbf{Cache-Passing Style.}
We can now transform our interpreter into cache-passing style.
We begin by changing the return type @Ans@ to a case class which contains a @VS@ set and
a cache.

@@

The skeleton of function @aeval@ largely remains the same.
Every place which has a value of @Ans@ is also replaced as direct pattern matching so
that we can use the fields in @Ans@ immediately.
The cache is used in a monotonic way: each function call to @aeval@ or @nd@ is
passed with the latest cache from the most recent continuation's result.

Upon entering @aeval@, we first determine whether the @out@ cache
contains some values for the current configuration and use them immediately (if present).
This corresponds to small-step AAM with a worklist: when we have seen a state,
we can simply discard the state and continue working through the rest of the worklist @todo@.

If, however, we have not hit the @out@ cache, we retrieve the values from the @in@ cache, and update the @out@ cache
with values from the @in@ cache. This retrieval from the @in@ cache may return an empty set of values if there is no such mapping for the configuration from the previous iteration.
In reality, we first conservatively assume that all computations can be diverged, and
if not, we update its values in the mapping after evaluation.
In terms of partial orders and lattices, it is the case that starts
the Kleene iteration from the bottom of the lattice.
